Hydra Battles and Cie

### 2020-12-25 doc/part-hydras.tex
minor modifications

@@

\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
\label{chap:ketonen}
\index{Maths!Ketonen-Solovay machinery}

@@

In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic. Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic. In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals. Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.

@@

We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

@@

\subsubsection{Limit Ordinals are Really Limits}
The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> limitb lambda  ->
  strict_lub (fun i => canonS lambda i) lambda.
\end{Coqsrc}

@@

\begin{exercise}\label{exo:simply-typed-canonseq}
Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$. Of course, you will have to prove the correctness of your function.
\end{exercise}

@@

>>>>

This lemma helps us

@@

\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
\label{chap:ketonen}
\index{Maths!Ketonen-Solovay machinery}

@@

In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic. Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic. In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals. Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.

@@

We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

@@

\subsubsection{Limit Ordinals are truly Limits}
The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> limitb lambda  ->
  strict_lub (canonS lambda) lambda.
\end{Coqsrc}

@@

\begin{exercise}\label{exo:simply-typed-canonseq}
Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$. Of course, you will have to prove the correctness of your function.

\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
The type \texttt{fuel}, an alternative 
to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
\index{Coq!Techniques!Giving fuel to a long computation}

@@

\begin{Coqsrc}
  Lemma m_lt : m small_h  t1< m big_h.

  Theorem Impossibility_std: False.
